---
title: "racir - Rapid A/Ci response analysis"
author: "Joseph R. Stinziano"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{racir - Rapid A/Ci response analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette explains how to process rapid A/Ci response (RACiR) data with the
'racir' package in R. It is written to be usable for beginners in R. Please
note that the functions in the package in its current form will only work with
Li-Cor 6800 (LI6800) data files (not the spreadsheet .xlsx output). Notes will
explain how to modify the function code for use with other types of data files.

After installing the package from CRAN with 'install.packages("racir")', load
the package :

```{r}
library(racir)
```

##Reading in Data
In the examples below I will use three built-in data files in the 'racir'
package: 'cal', 'poplar_1', and 'poplar_2'. These are LI6800 data files in
a tab-delimited format. These files can be called in using the code:

system.file("extdata", "FILENAME", package = "racir")

where FILENAME is one of the named files above.

The following example is to read in a data file from the LI6800. Note that
this is not restricted to RACiR data, but can be used for reading LI6800 files
in general.

The argument skiplines specifies the number of lines to skip in the tab
delimited file to reach the variable names (e.g. E, A, Ci, etc.). It defaults
to 53, but may vary depending on your machine and software version.

```{r}
data <- read_6800(system.file("extdata", "cal", package = "racir"), 
                  skiplines = 53)
```

##Checking Calibration Data
Since RACiR precludes matching the sample and reference gas analyzers, an empty
chamber RACiR needs to be run and used to correct the leaf sample data. First
however, it is necessary to ensure that only usable calibration data are
retained. The following code outputs three graphs: 'Check cutoffs' helps you to
visually determine where the unusable shoulder portions are of the RACiR.
'Confirm cutoffs' helps you to verify that your cutoff arguments are
appropriate. 'Calibration Fits' shows you how each polynomial fits the data.

```{r fig.height = 4, fig.width = 6}
racircalcheck(calfile = system.file("extdata", "cal", package = "racir"),
              skiplines = 53)
```

In the above example, there are shoulders in the A versus CO2_r response.
These shoulders represent less predictable chamber mixing at the beginning
and end of RACiRs. We can cut them out using the 'mincut' and 'maxcut'
arguments.

```{r fig.height = 4, fig.width = 6}
racircalcheck(calfile = system.file("extdata", "cal", package = "racir"),
              mincut = 350, maxcut = 780, skiplines = 53)
```

From the 'Check cutoffs' and 'Calibration Fits' graphs, we can see that the
cutoffs for CO2_r of 350 and 780 narrow the calibration file to a predictable
response.

##Correcting RACiR Files
The following code can be used to correct a single RACiR data file based on a
paired calibration file. This code should be used after running 'racircalcheck'
to determine appropriate cutoffs. The 'racircal' function is most appropriate
when you have one calibration file per leaf RACiR.

```{r fig.height = 4, fig.width = 6}
racircal(calfile = system.file("extdata", "cal", package = "racir"),
         mincut = 350, maxcut = 780,
         datafile = system.file("extdata", "poplar_2", package = "racir"),
         skiplines = 53)
```

If there is more than one leaf RACiR per calibration file, then a batch
analysis can be performed using the 'racircalbatch' function.

```{r fig.height = 4, fig.width = 6}
#Create a list of files
myfiles <- c(system.file("extdata", "poplar_1", package = "racir"),
             system.file("extdata", "poplar_2", package = "racir"))

#If all files in a folder pertain to one calibration file, you can use:
#myfiles = list.files()

racircalbatch(calfile = system.file("extdata", "cal", package = "racir"),
         mincut = 350, maxcut = 780,
         datafiles = myfiles, skiplines = 53)
```

##Compiling RACiR Files
Once the RACiR files have been processed, the files can be compiled into a .csv
file using the following code:

```{r eval = FALSE}
compile_racir(outputfile = "mydata.csv")
```
